A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-07-30 paper/paper.tex
paper: lots of progress on Our Construction; restructure

@@

\section{Our construction}
Our construction is highly dependent, with dependencies in inequality and equality proofs. It involves several involved details, some of them being visible only in the Coq formalization.

In tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames}, \ref{tab:faces}, \ref{tab:coh}, we give an informal, mathematical definition in Extensional Type Theory, which roughly corresponds to a level of reasoning ``à la set theory'' where equalities are thought as ``holding on the nose'' without having to mention them explicitly once they are justified. In section \ref{sec:itt}, we give the same mathematical definition in Intensional Type Theory, where transporting along coherence conditions is now made explicit. Adding this level of details helps to understand the specificities of coherence conditions.

\subsection*{Formal construction in ETT\label{sec:ett}}
% Abbreviated tables in this section
%
% The rule for implicit arguments for each table:
% show 1 explicit in first two parts
% show E + 1 explicit in the third part
\appendmask[bonak]{layer}[D]
\appendmask[bonak]{painting}[D]
\appendmask[bonak]{restrframe}[D]
\appendmask[bonak]{restrlayer}[D, d]
\appendmask[bonak]{restrpainting}[D, d]
\appendmask[bonak]{cohframe}[D]
\appendmask[bonak]{cohlayer}[D, d]
\appendmask[bonak]{cohpainting}[D, d]

@@

We now present the construction in \emph{extensional} type theory; i.e. in a type theory with the following reflection rule, where $=$ is propositional equality in some type and $\equiv$ is definitional equality~\cite{martinlof84}:
\begin{equation*}
  \seqr{}{\Gamma \vdash p: t = u}{\Gamma \vdash t \equiv u}
\end{equation*}

The definition is dispatched over tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames}, \ref{tab:faces} and~\ref{tab:coh}. Table \ref{tab:coind} describes a $\nu$-set in indexed form, as a stream, coinductively representing the limit of $n$-truncated $\nu$-sets. The $n$-truncated $\nu$-sets are themselves described in \ref{tab:core}. In such stream, the $n$th component is a type dependent over a $\fullframe$. The type $\fullframe$ is recursively defined in \ref{tab:frames}, using the auxiliary definitions of $\layer$ and $\painting$. The type $\layer$ and $\painting$ are dependent over $\restrf$, and these restrictions are defined on \ref{tab:faces}. These restrictions are defined using auxiliary definitions of $\restrl$ and $\restrc$.

@@

\section{Details on the mechanization\label{sec:mech}}
\subsection*{Formal construction in ITT\label{sec:itt}}
\renewcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\renewcommandx{\cohl}{\coh{layer}[][][][][][][][]}

We now present the definition in \emph{intensional} type theory~\citep{martinlof84,coq22}, making explicit the need for transport along coherence proofs.

>>>>

We are also interested in \emph{extensional} type theory; i.e. in a type theory with the following reflection rule, where $=$ is propositional equality in some type and $\equiv$ is definitional equality~\cite{martinlof84}:
\begin{equation*}
  \seqr{}{\Gamma \vdash p: t = u}{\Gamma \vdash t \equiv u}
\end{equation*}

@@

\section{Our construction}

From now on, we assume reasoning in type theory, first informally, then, stepwise, in fully formal type theory.

In informal type theory, which would roughly correspond to an extensional type theory leaving implicit trivial steps, our inductive construction is described by tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames} and \ref{tab:faces}. However, we went to the point of fully formalizing it in the intentional type theory of the Coq proof assistant. Going to such level of details requires to address the following points:
\begin{itemize}
  \item Verification that the construction is well-typed: for instance, the typing of $\restrl$, given in Table \ref{tab:faces}, requires a non-trivial coherence condition, given on \ref{tab:coh}, which has to be proved simultaneously with the rest of the construction. This proof also relies on type isomorphisms which have to be made explicit. All this is discussed in Section ...
  \item The well-foundedness of the induction requires a special termination evidence which will be discussed in Section ...
  \item Technically, even if trivial from the mathematical, the components of the construction have side conditions on the range of parameters of the constructions. In type theory, this typically corresponds to be dependent on proofs of inequalities on natural numbers. We discuss how we deal with such dependencies in Section ...
  \item The simulation of extensional type theory in intensional type theory requires to make explicit equality reasoning, leading to types of the construction becoming dependent on proofs of equalities. This will be discussed in Section ...
  \item Technically, a number of groupoid properties of equality are also used which will be discussed in Section ...
\end{itemize}

But first, let us give details and intuition on the informal construction.

\subsection*{The informal construction\label{sec:ett}}
% Abbreviated tables in this section
%
% The rule for implicit arguments for each table:
% show 1 explicit in first two parts
% show E + 1 explicit in the third part
\appendmask[bonak]{layer}[D]
\appendmask[bonak]{painting}[D]
\appendmask[bonak]{restrframe}[D]
\appendmask[bonak]{restrlayer}[D, d]
\appendmask[bonak]{restrpainting}[D, d]
\appendmask[bonak]{cohframe}[D]
\appendmask[bonak]{cohlayer}[D, d]
\appendmask[bonak]{cohpainting}[D, d]

@@

The definition is thus dispatched over tables \ref{tab:coind}, \ref{tab:core}, \ref{tab:frames} and \ref{tab:faces} with~\ref{tab:coh} showing the proof of a needed coherence condition. Table \ref{tab:coind} describes a $\nu$-set in indexed form, as a stream, coinductively representing the limit of $n$-truncated $\nu$-sets. The $n$-truncated $\nu$-sets are themselves described in \ref{tab:core}. In such stream, the $n$-th component is a type dependent over a $\fullframe$. The type $\fullframe$ is recursively defined in \ref{tab:frames}, using the auxiliary definitions of $\layer$ and $\painting$. The type $\layer$ and $\painting$ are dependent over $\restrf$, and these restrictions are defined on \ref{tab:faces}. These restrictions are defined using auxiliary definitions of $\restrl$ and $\restrc$.

@@

\subsection*{From extensional to intentional type theory: explicitation of the equality proofs\label{sec:itt}}
\renewcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\renewcommandx{\cohl}{\coh{layer}[][][][][][][][]}
